Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2017-11-23 articles/formalism/sections/real_space_formalism.tex
revised discussion in sec 3.1 and 3.2, yet to complete

Specifically note that $\alpha_{qe}, ~\beta_{qe} ~\&~ \gamma_{qe}$ denote the Euler angles which rotate the local cartesian system at the location $\hat{n}_q$ where the Stokes field is measured to the cartesian system at the location $\hat{n}_e$ where the scalar fields are evaluated: $\hat{n}_q \xrightarrow{\mathcal{R}(\alpha_{qe},\beta_{qe},\gamma_{qe})} \hat{n}_e$. The Euler angles for the inverse rotation are given by the Euler angles $\alpha_{eq}=-\gamma_{qe}$, $\beta_{eq} = -\beta_{qe}$ and  $\gamma_{eq} =-\alpha_{qe}$\cite{varshalovich}. Since the real space kernels will only depend on the cosine of the Euler angle $\beta$, they are immune to its change of sign. However the ordering of indices  on the Euler angle $\alpha$ and $\gamma$  is important and leads to two different interpretations of the real space kernels.

\textit{Convolution:} Now working from the position of the pixel ``e'' where the scalar modes need to be evaluated, the Euler angles to the surrounding pixels are: $\alpha_{eq}, \beta_{eq},\gamma_{eq}$. One can now think of the value of scalar fields $E$ \& $B$ at the central pixel ``e" as resulting from the convolution over the measured Stokes $Q$ \& $U$ parameters in the surrounding pixels ``q''. This convolution is defined in terms of the Euler angle $\gamma_{eq}$ and is given by the following equation,
%
\beq \label{eq:qu2eb_convolution_explicit}
\bmat E_e \\ B_e  \emat =- \sum_{q=1}^{N_{\rm pix}}{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \gamma_{eq}) & -\sin(2\gamma_{eq})\\  \sin(2 \gamma_{eq})  & \cos(2 \gamma_{eq}) \emat  \bmat Q_q \\ U_q  \emat \Delta \Omega\,,
\eeq
%
where $\Delta \Omega$ denotes the pixel area and all the symbols have their usual meaning. 
%\revisit{This has an elegant interpretation: to derive the E and/or B field at any given position we need to find the cosine quadrupole transform and the sine quadrupole transform of the Stokes Q \& U parameters on circles around this position, weigh the transform by the value of the function $f(\beta,\ell_{\rm min},\ell_{\rm max})$, $\beta$ being the radius of the circle and sum up the results with appropriate signs, to construct the respective scalar fields.} 
The above equation can be expressed more concisely as follows,
%
\begin{subequations} \label{eq:qu2eb_simple}
\beqry 
[E + iB](\hat{n}_e) &=& - \Delta \Omega \sum_{q=1}^{N_{\rm pix}} \left(\sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell+2)!}} P_{\ell}^{2}(\beta_{eq}) \right) {\Bigg( e^{i2 \gamma_{eq}}   {}_{+2}X (\hat{n}_q) \Bigg)} \,, \label{eq:qu2eb_physical}\\
&=& - \left( \left[ \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \sqrt{\frac{2 \ell+1}{4 \pi}}Y_{\ell 2}(\beta,\gamma) \right]  \circ {}_{+2}X \right) \,, \\
&=& - \Bigg\lbrace \mathcal{M}^* \circ {}_{+2}X \Bigg\rbrace(\hat{n}_0)\,, \label{eq:qu2eb_convolution} 
\eeqry
\end{subequations}
%
where $\circ$ denotes a convolution and the spherical harmonic functions denote the rotated functions such that the pole of the function coincides with the direction $\hat{n}_e$. %and the reference zero for the azimuthal angle is the local longitude $\phi_0$.

\textit{Radiating:} Working from the position ``q'' where the Stokes parameters are known, the Euler angles to the surrounding pixels are: $\alpha_{qe}, \beta_{qe},\gamma_{qe}$. The E \& B mode maps resulting from the Stokes Q \& U parameters specified at location $\hat{n}_q$ is given by the following expression,
%
\beq \label{eq:qu2eb_radiation_explicit}
\bmat E_e \\ B_e  \emat_q =- {{}_{\mm}f}(\beta_{qe},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{qe}) & \sin(2\alpha_{qe})\\  -\sin(2 \alpha_{qe})  & \cos(2 \alpha_{qe}) \emat  \bmat Q_q \\ U_q  \emat \Delta \Omega \,, 
\eeq
%
and the total map of E \& B modes can be simply evaluated by summing over the contribution from the Stokes parameters at each location $\hat{n}_q$ : $\bar{S} = \sum_{q=1}^{N_{\rm pix}} \bar{S}_q$.
%$\bmat E_e \\ B_e  \emat = \sum_{q=1}^{N_{\rm pix}} \bmat E_e \\ B_e  \emat_q$.

@@

We do not provide the explicit derivation here, since the procedure is nearly identical to that discussed in the previous section. The inverse operator is given by the following expression,

@@

{\bar O}^{-1}=-\bmat \mathcal{M}_{r} & -\mathcal{M}_{i} \\  \mathcal{M}_{i}  & \mathcal{M}_{r} \emat_{2 N_{\rm pix} \times 2 N_{pix}} =-{{}_{\mm}f}(\beta_{ij},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{ij}) & -\sin(2\alpha_{ij})\\  \sin(2 \alpha_{ij})  & \cos(2 \alpha_{ij}) \emat \,.

@@

%where all the symbols have the same meaning as discussed in \sec{sec:qu2eb}.
Note that the kernel is different by a mere change in sign on the off-diagonals of the block matrix as compared to \eq{eq:op_qu2eb}. We can evaluate the Stokes parameters  $Q$ \& $U$ from the scalar fields $E$ \& $B$ by evaluating the following expression,

>>>>

Specifically note that $\alpha_{qe}, ~\beta_{qe} ~\&~ \gamma_{qe}$ denote the Euler angles which rotate the local cartesian system at the location $\hat{n}_q$ where the Stokes field is measured to the cartesian system at the location $\hat{n}_e$ where the scalar fields are evaluated: $\hat{n}_q \xrightarrow{\mathcal{R}(\alpha_{qe},\beta_{qe},\gamma_{qe})} \hat{n}_e$. The Euler angles for the inverse rotations are given by the Euler angles $\alpha_{eq}=-\gamma_{qe}$, $\beta_{eq} = -\beta_{qe}$ and  $\gamma_{eq} =-\alpha_{qe}$\cite{varshalovich}. Since the kernels only depend on the cosine of the Euler angle $\beta$, they are immune to the change in its sign. More importantly, the ordering of indices  on the Euler angle $\alpha$ and $\gamma$  is important and leads to two different interpretations of the real space kernels.

\textit{Convolution:} Working from the frame of the pixel ``e'' where the scalar modes need to be evaluated, the Euler angles to the surrounding pixels are: $\alpha_{eq}, \beta_{eq},\gamma_{eq}$. In terms of these Euler angles, the value of scalar fields $E$ \& $B$ at the central pixel ``e" can be evaluated by convolving over the measured Stokes $Q$ \& $U$ parameters in the surrounding pixels marked by indices ``q''. Specifically, the convolution is fully defined in terms of the Euler angles $\beta_{eq}~ \& ~\gamma_{eq}$ and is given by the following equation,
%
\beq \label{eq:qu2eb_convolution_explicit}
\bmat E_e \\ B_e  \emat =- \sum_{q=1}^{N_{\rm pix}}{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \gamma_{eq}) & -\sin(2\gamma_{eq})\\  \sin(2 \gamma_{eq})  & \cos(2 \gamma_{eq}) \emat  \bmat Q_q \\ U_q  \emat \Delta \Omega\,,
\eeq
%
where $\Delta \Omega$ denotes the pixel area and all the symbols have their usual meaning. 
%\revisit{This has an elegant interpretation: to derive the E and/or B field at any given position we need to find the cosine quadrupole transform and the sine quadrupole transform of the Stokes Q \& U parameters on circles around this position, weigh the transform by the value of the function $f(\beta,\ell_{\rm min},\ell_{\rm max})$, $\beta$ being the radius of the circle and sum up the results with appropriate signs, to construct the respective scalar fields.} 
When written in terms of the complex spin-0 field $[E + iB]$ and the spin-2 field ${}_{+2}X$, the above equation can be expressed more concisely as follows,
%
\begin{subequations} \label{q:qu2eb_convolution_concise}
\beqry 
%[E + iB](\hat{n}_e) &=& - \Delta \Omega \sum_{q=1}^{N_{\rm pix}} \left(\sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell+2)!}} P_{\ell}^{2}(\beta_{eq}) \right) {\Bigg( e^{i2 \gamma_{eq}}   {}_{+2}X (\hat{n}_q) \Bigg)} \,, \label{eq:qu2eb_physical}\\
[E + iB](\hat{n}_e) &=& \Bigg\lbrace \left[ - \Delta \Omega \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \sqrt{\frac{2 \ell+1}{4 \pi}}Y_{\ell 2}(\beta_{eq},\gamma_{eq}) \right]  \circ {}_{+2}X \Bigg\rbrace(\hat{n}_e) \,, \\
&=& \Bigg\lbrace \mathcal{M}_{Beam} \circ {}_{+2}X \Bigg\rbrace(\hat{n}_e)\,, \label{eq:qu2eb_convolution} 
\eeqry
\end{subequations}
%
where $\circ$ denotes the convolution and the spherical harmonic functions denote the rotated functions such that the pole of the function coincides with the direction $\hat{n}_e$. %and the reference zero for the azimuthal angle is the local longitude $\phi_0$.

\textit{Radiating:} On the contrary, while working from the frame of the pixel ``q'', the Euler angles to the surrounding pixels are: $\alpha_{qe}, \beta_{qe},\gamma_{qe}$. The E \& B mode maps resulting from the Stokes Q \& U parameters specified at location $\hat{n}_q$ is given by the following expression,
%
\beq  \label{eq:qu2eb_radiation_explicit}
\bmat E_e \\ B_e  \emat_{q} =- {{}_{\mm}f}(\beta_{qe},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{q e}) & \sin(2\alpha_{q e})\\  -\sin(2 \alpha_{q e})  & \cos(2 \alpha_{q e}) \emat  \bmat Q_{q} \\ U_{q}  \emat \Delta \Omega \,.
\eeq
%
The total map of E \& B modes can be simply evaluated by summing over the contribution from the Stokes parameters at each location $\hat{n}_q$ : $\bar{S} = \sum_{q=1}^{N_{\rm pix}} \bar{S}_q$. The above expression can again be cast in a more concise form as follows,
%
\begin{subequations} \label{eq:qu2eb_radiation_concise}
\beqry 
\left[E + iB\right] &=& -\Delta \Omega  \sum_{q=1}^{N_{\rm pix}} \Big[ {}_{+2}X(\hat{n}_{q}) e^{-i2\alpha_{q e}} \Big]  {{}_{\mm}f}(\beta_{q e}) \,, \\
&=&    \sum_{q=1}^{N_{\rm pix}} \Bigg\lbrace {}_{+2}X(\hat{n}_{q}) \cdot \left[ - \Delta \Omega \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \sqrt{\frac{2 \ell+1}{4 \pi}}Y^*_{\ell 2}(\beta_{qe},\alpha_{qe}) \right] \Bigg\rbrace \,, \\
&=& \sum_{q=1}^{N_{\rm pix}} {}_{+2}X(\hat{n}_{q}) \cdot  \mathcal{M}_{Green}(\hat{n}_{q}) \,,
\eeqry
\end{subequations}
%
where ``$\cdot$'' denotes a simple scalar multiplication. \revisit{$\mathcal{M}_{Green}(\hat{n}_{q})$ is response of the real space operator $\bar{O}$ when operated on a Stokes field: $\delta(\hat{n}-\hat{n}_q) + i \delta(\hat{n}-\hat{n}_q)$, it can be thought of as the Green's function of the real space operator.} \comment{Is this interpretation correct ? Is the complex delta function a spin 2 function ?}

@@

We do not provide the explicit derivation here, since the procedure is identical to that discussed in the previous section. The inverse operator is given by the following expression,

@@

{\bar O}^{-1}=-\bmat \mathcal{M}_{r} & -\mathcal{M}_{i} \\  \mathcal{M}_{i}  & \mathcal{M}_{r} \emat_{2 N_{\rm pix} \times 2 N_{pix}} =-{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{eq}) & -\sin(2\alpha_{eq})\\  \sin(2 \alpha_{eq})  & \cos(2 \alpha_{eq}) \emat \,.

@@

where all the symbols have the same meaning as discussed in \sec{sec:qu2eb}. Note that the kernel in the above equation differs from the one in \eq{eq:op_qu2eb} by a change in sign on the off-diagonals of the block matrix and the pixel indexing is reversed $ e \leftrightarrow q$. We can evaluate the Stokes parameters  $Q$ \& $U$ from the scalar fields $E$ \& $B$ by evaluating the following expression,
